\documentclass[oneside, article]{memoir}
% \input{../packages}
\input{../../../work/packages}
\input{../../../work/packagesMemoir}

\title{Moving}
\author{}

\begin{document}
\maketitle

\part{Apartment shifting and hunting}
\chapter{Shifting apartments}
\section{Cancel services}
\subitem Time-Warner internet.
\subitem City of Austin. Lease sometimes requires you to disconnect 3 days later.

\section{ Acquire services and utilities}
City of Austin electricity. (website or on 4949400).

Visit website, set up facility for receiving bills by email.

\subsection{Internet service providers}
Time warner, after the initial Earthlink promotion, is expensive.

ATT DSL costs 30\$ for 1.5 to 3 mbps when connected with iphone account. nAgarAjan pays 40\$ for 6mbps.

\section{Mail service}
\subsection{Set up address forwarding}
Attach name to USPS mailbox

\subitem print USPS form online, give it to carrier or take it to post office.

\subitem with apartment management.

\subsection{ Fix address with civic authorities and insurers}
\subitem TX DPS (extra money, not eligible for online address change).
\subitem TX Travis county tax office (no easy online change form).
\subitem AAA.

\subsection{Fix credit card addresses}
\subitem Amazon chase card: online.
\subitem UFCU credit card.

\subsection{ Fix bank contact addresses}
\subitem BoA (by going in person)
\subitem UFCU (go in person, or phone)

\subsection{Fix address with universities}
\subitem UMass, on spire website.
\subitem UTexas, on website and website 2. (CS dept address used.)

\subsection{With shopping websites}
amazon, google products, ebay/ paypal.

\section{Clean old apartment}
\subitem Defrost refridgerator.

\subitem Empty everything.

\subitem Remove all stains.

\subitem Clean fan, sweep floorboard.

\subitem Cleaning tiles and tub: Rolando recommended multiple applications of tilex. Internet recommends scrubbing with baking soda and hydrogen peroxide.

\subitem Cleaning ovens: Rolando recommended 'Easy Off Oven cleaner' with an abrasive called s(t)eel wool.

\subitem May be use professional cleaners: see craigslist.

\section{Clear and surrender old apartment}
\subitem Move luggage.

\subitem Return apartment, mail, laundry keys, parking stickers.

\section{Move into new apartment}
Photograph the place while moving in and moving out.

Pay rent.

Note defects in the condition form.

Fix bill pay.

\chapter{Apartment hunting}
Use an excel sheet (see SFO moving by Shruthi).

\section{Finding good apartments}
http://www.housingmaps.com searches craigslist ads on google maps. sulekha.com is useful for finding Indian apartment vacancy postings.

Note map with 20 minute walking radius of the department. Just search for "apartment" in google maps; see the apartmentreviews map. Drive around and note "leasing" signs. Start searching for an apartment around 1.5 months before the planned move-in, when people will have given move-out notices. Contact apartment finders at the suitable time.

\section{Good and close studios and 1 BHK}
\subitem Matthew properties: Duval, Elmwood: 512 454 0099. Rolando: 512 731 6799. Around 40 efficiencies and 1 BHK. Prompt management. They do small leases. Determine apt price: RK pays \$450, 495 sq ft. Slight bug infestation. 525\$ for \#207 above RK, with 350 deposit.
\subsubitem Efficiency 420 sq ft; Must move in on 1st; Good AC; 550\$. Equivalent Apt near the stair available for 535\$; must move in on 7th.
\subsubitem 1BHK for \$575.

\subitem Century apartments Dean Keeton: 512 327 6968. Negatives: Management slow to respond to maintainence requests; Noise from Dean Keeton St. 545\$. Around 8 units.

\subitem Campus crossings apt: 512 322 0512. Management may not do a thorough job in fixing problems. Very tiny: like a jail cell, an electric range is provided separately. Unfurnished. 545\$. Around 25 to 30 units.

\subitem Chaparosa: Slightly more spacious than Century or campus crossings. 525\$. Bug infestation, but this can be avoided with good maintenance. 512-481-8600.

\subitem Leon apartments: 2207 Leon St, Austin, TX 78705 512-474-2215 Good reviews, glass on parking lot, slow maintenance, 475 sq feet.

\subitem 2302 leon st. zip code 78705. 512 476 6616, 512 497 8282, Pat 512 363 7294. Ricardo's apartment complex. 495\$. Friendly management. No disposal. 300 sq feet.

\subitem 305 apartments: good reviews; 570\$.

\subitem ElDorado: smaller than El Campo; 600\$; Ardavan had no complaints; . ElCampo: 39th and Speedway; Ardavan is getting a special rate of 600.

\chapter{Checks and negotiations}
\section{Things to check}
\subitem Check facility provided to control AC/ heater. If electricity and water are submetered, find out how much the bill is in winter and in summer.
\subitem Check crime rate. Check apartment reviews online.
\subitem Subletting policy.

\section{Negotiation}
Note that you can negotiate with the management over rent.
\subitem Suggest a lower rent for signing a 1 year lease, rather than a 6 month lease.
\subitem Ardavan got a bigger apartment unit managed by the same company for the same rent upon negotiation.



\part{Travel protocol}
\chapter{Pre travel preparations}
\section{Prepare home for departure}
\subitem Ensure rent, if due during journey, is paid.
\subitem Throw trash.
\subitem Dispose food which will get spoiled.
\subitem Set heater to 48 F.

Replenish cell phone and calling card charge. Set 'travelling abroad' voice mail.

\section{Prepare documents}
\subitem Address and contact sheets.
\subitem Domestic flight tickets.
\subitem Scholastic emergency services and health insurance cards.
\subitem ID: Passport or drivers license.
\subitem Copies of important documents.

Prepare flash drive.

\subsection{For international travel}
Gather important documents. (Passport/ visa/ I-94, I-20, academic records, tickets, customs info.)
\subitem Ensure that I-20 has signature which is at most 6 (to 12?) months old.

If coming to USA, include:
\subitem Letter of Good Standing, also called Letter of Enrollment (obtained at the Registrars Office)
\subitem Offical UT Transcript (obtained at the Registrarar's Office - Main Building 1)
\subitem Proof of financial support (a personal bank statement, Research Assistant/Teaching Assistant verification letter that includes salary and tuition payment details.

Prepare address/ phone sheet [with calling card information for places visited, address information at source and destination, travel agent and airline numbers].

Prepare copies of travel documents. [Passport, Visa, I-94, I-20, address/ phone sheet, tickets, customs info]


\section{Gather travel details}
Acquire milage credits.

Determine weight limits.

Check flight departure time/ status: Use the airline confirmation codes to check flight status, not just the travel booking agent website. Eg: In May 2010, Vayama did not send me a notification when the frankfurt - bengaluru flight was cancelled.

\section{Financial arrangements}
Inform credit / debit card companies of travel plans to ensure that credit cards are not blocked.
\subitem BoA debit card center: 18778335617, customer care : 18004321000, India prefix 000117 or use 1.925.681.7600.

\subsection{For international travel}
Acquire required currency.

Acquire health insurance if needed.


\section{Pack sustenance items}
\subsection{To be worn}
\subsubitem Socks. Shirt, underwear, suitable pants, kerchiefs.
\subsubitem A pair of boots.
\subsubitem 1 sweatshirt, 1 jacket.
\subsubitem Wallet. Driver's license. Credit cards, a check leaf.

Label/ tag the luggages.

Wear coat, shoes: airports may be cold.


\section{Cabin luggage}
\subsubitem 2 or 3 pairs of Socks.
\subsubitem quick dry Swimming trunk, an underwear.
\subsubitem Gloves, warm hood.
\subsubitem Tooth brush and paste, tongue scraper.
\subsubitem GPS. Camera, if necessary.
\subsubitem Medicines: Petroleum jelly.
\subsubitem Chewing gum.

Pack work items:
\subitem Laptop, its charger, microphone/ earphone.
\subitem Cell phone, phone charger.
\subitem Papers and books to read.
\subitem Pens, scratch pad, small notebook.

\subsection{To be carried as luggage}
\subsubitem 3 pants with kerchiefs. If visiting cold weather, carry or wear warm pants.
\subsubitem Water and windproof pant shell, jacket shell.
\subsubitem 1 sweatshirt, one of them warm.
\subsubitem Underwear. Also, Long underwear for cold weather.
\subsubitem Slippers.
\subsubitem 2 or 3 pairs of Socks.
\subsubitem Essential sustenance kit, including scissors, tape, nail cutter, shaving blade, sewing thread, twig for wearing caste mark.

\subitem Take emergency medicine: pain relief, acidity, allergy.
\subitem Take soap strips, toothbrush, comb.

\section{During Travel}
Check-in, fly.

You can get boarding pass for flights of other airlines at the gate of one airline - generally, there are agreements amongst airlines.

While boarding the flight, at each gate, confirm transfer of luggage, meal, seating preferences. American airways had not ensured vegetarian meals for me during mid2010 visit to bhArata.

\section{International travel}
While travelling from USA to outside, ensure that I-94 form is taken away.

\section{After travel}
Update glacier with travel dates.

\end{document}
