\documentclass[oneside, article]{memoir}
\input{../../../work/packages}
\input{../../../work/packagesMemoir}
\title{sAmagri-yojanA-sUtraM}
\author{vishvAs}

\begin{document}
\maketitle
\part{parichayaH}


\part{kArya-sAmagrayaH}
\chapter{Computer}
See ergonomics section in shArIrika-kriyA-sUtra.

\section{krENa-tantraM}
When new models are released, prices of older models drop. Competition also lowers prices. Wait for these.

Refurbished items are cheaper.

Use deals websites like onsale.com. 

\chapter{Chairs}
\section{Cushioning, support}
The base and the back-rest should be padded - mesh support is uncomfortable. For the comfort of the buttock, exterior should be a finely knit cloth - a deficiency here can be filled by spreading a towel or a bed-sheet.

Mesh backrests are usually unsatisfactory.

\section{Adjustability}
The following should be precisely and firmly adjustable: Position on the floor (using wheels), face direction (by swiveling), seat depth (and, potentially footrest height), back-rest height,  back-rest tilt angle.

\subsection{Effort}
If they are adjustable with low effort - like at the push of a lever, the chair can easily be adjusted to varying demands of the body. This should atleast be the case with frequently adjusted components like back-rest angle and seat depth.

The wheels should move freely even when no one is sitting on it.

\section{Armrest}
Armrests are generally not necessary, as wrists/ elbows can be rested on the table during use.

\section{Footrest}
The chair may possess an adjustable footrest/ foot ring- but that should be wide and firm enough to be useful. The footring should be sufficiently wide.

Otherwise, a foot stool may be used.

\chapter{Tables}
Buy a wide table which can hold both the computer and writing/ reading material.

Height adjustability is important, but can be compromised on if the chair used with the table is sufficiently height adjustable, and has a good footrest.

Ideally, the table should be adjustable so as to allow working while standing up. Eg: Ikea FREDRIK work station (201.159.91).


\part{gRRiha-visheShAH}
\chapter{Conversation chairs}
Buy ottomans which also serve as storage bins.

\chapter{Storage}
Buy plastic tray trolleys with wheels.

Buy collapsible cloth + cardboard shelfs which can be hung from a rod or a mobile rod with wheels.

\chapter{Transportation}
Buy luggage with wheels - avoid things you need to carry on your shoulder.

\end{document}
